\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual Multi-Loop Integrands

### 2023-09-21 notes/2loopNLSM.tex
Update 2loopNLSM.tex

It is tempting to try to match the pion four-point two-loop integrand to ZM
on the cuts, but given that the cuts only probe the on-shell four-point
amplitude, which is either subtle or vanishes for ZM, it is better to
compare the off-shell numerators directly.
The numerators of the two basis graphs are
\be
\ZMBasisDoubleBox =\left[\begin{split}
  \langle (l_1+k_4) l_1 \rangle\langle l_1l_2 \rangle\langle l_2 (l_2+k_1) \rangle  \langle (l_2+k_1)(l_2+k_1+k_2) \rangle 
\\
\langle (l_1+k_3+k_4)(l_2+k_1+k_2) \rangle\langle (l_1+k_3+k_4)(l_1+k_4) \rangle
\end{split}\right]
\ee
\be
\ZMBasisPentaTriangle =\left[\begin{split}
  \langle (l_1+k_4) l_1 \rangle\langle l_1l_2 \rangle\langle l_2 (l_2+k_1) \rangle  \langle (l_2+k_1)(l_1-k_1) \rangle 
\\
\langle (l_1+k_3+k_4)(l_1-k_1) \rangle\langle (l_1+k_3+k_4)(l_1+k_4) \rangle
\end{split}\right]
\ee
where $\langle ab\rangle \equiv p_a^\mu \varepsilon_{\mu\nu} p_b^\nu$ is the color-stripped ZM vertex.
As local functions, the
numerators never suffer from any of the subtleties of on-shell 2D
kinematics.
We introduce a proportionality constant $z$ between the numerators of the two theories, $N_\text{NLSM} \vert_{2D} = z ~ N_\text{ZM}$, because we are only interested in how the kinematical structure of $N_\text{ZM}$ can be used to eliminate gauge freedom in $N_\text{NSLM}$.\footnote{Another reason for introducing the parameter $z$ is because the theories are believed to differ at loop level even though they are dual classically in 2D \cite{Nappi:1979ig}.  Once all of the constraints from this section are imposed, $z$ is fixed to $216/565$ where the four-point pion tree amplitude is normalized to $-k_1\cdot k_2$ and the ZM vertex is normalized to $k_1^\mu \varepsilon_{\mu\nu}k_2^\nu$.}
Mechanically, the pion numerator is matched
to ZM by first taking every possible assignment of right and left
movers for the external particles and then restricting the pion
numerator to 2D.  The loop momenta are restricted to 2D but left
off-shell.  After performing the 2D matching, there are only 365
parameters of generalized gauge freedom.

@@

\section{Conclusions and Outlook}\label{conclusions}

There is overwhelming evidence that YM and NLSM obey color-kinematics
duality at loop level, even though naive Lagrangians for these theories
do not manifest the duality.  At tree level, the duality can be
manifested by breaking manifest Bose symmetry at the Lagrangian level,
either by explicitly picking different gluon states (as in SDYM) or by
picking states that only restore Bose symmetry in the final amplitude
(as in NLSM).  All of these theories can be encapsulated in the
semi-abelian YM theory presented in \cref{eq:semiYM} which obscures
the Bose symmetry of the external states due to the presence of the
two gauge fields $A_\mu$ and $\bar{A}_\mu$.  The kinetic mixing
between the two gauge fields permits globally color-dual 1-loop
integrands, but special care must be taken so that the integrands
produced from semi-abelian YM correspond to SDYM or NLSM.  Due to the
kinetic mixing that stems from the broken Bose symmetry, it is
impossible to construct a two-loop integrand for semi-abelian YM theory,
that is, the theory is one-loop exact.  Obtaining a globally color-dual
integrand beyond one-loop then requires sacrificing more than just Bose
symmetry and in this paper it is argued that off-shell locality must
be abandoned.\footnote{Locality is relaxed in \cite{Mogull:2015adi} in
  order to obtain color-dual two-loop integrands.}

Kinematic ``numerators'' that are rational functions of the kinematics
appear unnatural from the perspective of point-like quantum field
theories. After all, operators that produce rational functions of
kinematics must be non-local. Thus, in order to identify valid color-dual representation of
Yang-Mills beyond tree-level, we must relax the constraint of
\textit{off-shell locality}, and realize locality only in the limit of
\textit{on-shell} kinematics. Below we provide an overview of what one
might expect for such a construction, and provide some simple examples
at tree-level that manifest this structure.

\subsection{Non-local construction of scattering amplitudes}
\label{nonLocalScattering}
As an exemplar of an off-shell non-local construction, consider the simple four-point example of a color-dual
representation of Yang-Mills theory. We can define a functional numerator, $N_{(12|34)}^{\text{YM}}$, as follows, 
\begin{equation}
N_{(12|34)}^{\text{YM}} = \frac{t_8F^4}{3} \frac{s_{13}-s_{12}}{s_{23}{s_{13}}}
\end{equation}
where $N_s = N_{(12|34)}$, $N_t = N_{(14|23)}$ and $N_u = N_{(13|42)}$. By construction, this functionally symmetric numerator is
anti-symmetric and obeys the Jacobi identity:
\begin{equation}
N_{(12|34)}^{\text{YM}} + N_{(13|42)}^{\text{YM}} +N_{(14|23)}^{\text{YM}} = 0
\end{equation}
However, it must also factorize to kinematics that are consistent with
the local Feynman rules of \cref{eq:ym-lag}. Evaluating the residue on
the $s=s_{12} \to 0$ pole we find:
\begin{align}
\text{Res}\left[\frac{ N_{s}^{\text{YM}} }{s}\right]^{s=0}&= \frac{2}{3}A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\\
\text{Res}\left[\frac{N_{t}^{\text{YM}} }{t}\right]^{s=0}&= -\frac{1}{3}A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\\
\text{Res}\left[\frac{N_{u}^{\text{YM}} }{u}\right]^{s=0}&= -\frac{1}{3} A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\end{align}
with $t=s_{23}$ and $u=s_{13}$.
Plugging these into a cubic graph representation of Yang-Mills, the
tree level version of \cref{eq:gen-amp}, we find:
\begin{equation}
  \text{Res}\left[\mathcal{A}_4^{\text{YM}}\right]^{s=0}
  = \frac{1}{3} A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} (2c_s-c_t-c_u)
  =  \mathcal{A}_{(12|l)}^{\text{YM}}\mathcal{A}_{(-l|34)}^{\text{YM}} 
\end{equation}
where we have applied the color structure Jacobi identity,
$c_s+c_t+c_u=0$. In light of our findings in \cref{chefsKiss},  While we can absolutely apply a generalized gauge
transformation \cite{BCJ} to restore field theoretic locality to the
cubic numerators, it is maybe aesthetic
choice.

Moreover, by abandoning traditional notions of off-shell
locality, we incidentally gained enough freedom to massage the
color-dual numerators into a form that is manifestly gauge invariant
for all particles. As an organizational principle, pulling out overall
gauge-invariant factors comes with the added advantage of simplifying
a loop-level construction of color-dual numerators for Yang-Mills.

Much of the complexity for $D$-dimensional vector theories at
loop-level arises due to the mixing between external polarization and
internal loop momenta. As noted, one advantage of expanding ansatz
construction towards functions of rational kinematics is the prospect
of stripping out overall factors of gauge invariant tensors. Rather
than building an ansatz from the irreducible scalar products of
\cref{eq:ym-basis},
\begin{equation}
N^{\text{YM}}_{m,L} = \sum_i \mathcal{O}_i \mathcal{R}^{(i)}_{m,L}
\end{equation}
where $\mathcal{O}_i$ are the basis tensors of
\cite{Bern:2017tuc,Carrasco:2019yyn}, and $ \mathcal{R}^{(i)}_{m,L}$ is a
rational function of scalar products.
\subsection{Future Directions}\label{sec:Discussion}

>>>>

\begin{figure}[t]
  \begin{center}
	\PhysicalCutOne{}{}{}{} \PhysicalCutTwo{}{}{}{}
  \end{center}
  \caption{The two physical cut topologies shared by NLSM, sGal, BI, and DBIVA}
  \label{fig:emu}
\end{figure}
It is tempting to try to match the pion four-point two-loop integrand to ZM
on the cuts, but given that the cuts only probe the on-shell four-point
amplitude, which is either subtle or vanishes for ZM, it is better to
compare the off-shell numerators directly.
The numerators of the two basis graphs are
\be
\ZMBasisDoubleBox =\left[\begin{split}
  \langle (l_1+k_4) l_1 \rangle\langle l_1l_2 \rangle\langle l_2 (l_2+k_1) \rangle  \langle (l_2+k_1)(l_2+k_1+k_2) \rangle 
\\
\langle (l_1+k_3+k_4)(l_2+k_1+k_2) \rangle\langle (l_1+k_3+k_4)(l_1+k_4) \rangle
\end{split}\right]
\ee
\be
\ZMBasisPentaTriangle =\left[\begin{split}
  \langle (l_1+k_4) l_1 \rangle\langle l_1l_2 \rangle\langle l_2 (l_2+k_1) \rangle  \langle (l_2+k_1)(l_1-k_1) \rangle 
\\
\langle (l_1+k_3+k_4)(l_1-k_1) \rangle\langle (l_1+k_3+k_4)(l_1+k_4) \rangle
\end{split}\right]
\ee
where $\langle ab\rangle \equiv p_a^\mu \varepsilon_{\mu\nu} p_b^\nu$ is the color-stripped ZM vertex.
As local functions, the
numerators never suffer from any of the subtleties of on-shell 2D
kinematics.
We introduce a proportionality constant $z$ between the numerators of the two theories, $N_\text{NLSM} \vert_{2D} = z ~ N_\text{ZM}$, because we are only interested in how the kinematical structure of $N_\text{ZM}$ can be used to eliminate gauge freedom in $N_\text{NSLM}$.\footnote{Another reason for introducing the parameter $z$ is because the theories are believed to differ at loop level even though they are dual classically in 2D \cite{Nappi:1979ig}.  Once all of the constraints from this section are imposed, $z$ is fixed to $216/565$ where the four-point pion tree amplitude is normalized to $-k_1\cdot k_2$ and the ZM vertex is normalized to $k_1^\mu \varepsilon_{\mu\nu}k_2^\nu$.}
Mechanically, the pion numerator is matched
to ZM by first taking every possible assignment of right and left
movers for the external particles and then restricting the pion
numerator to 2D.  The loop momenta are restricted to 2D but left
off-shell.  After performing the 2D matching, there are only 365
parameters of generalized gauge freedom.

@@

\section{Conclusions and Outlook}\label{conclusions}
In summary, we have found that there is a tension between color-kinematics duality and an off-shell local construction of the two-loop four-point of the pure Yang-Mills integrand in $D$-dimensions.
At tree level, we demonstrated that the duality can be
manifested by breaking manifest Bose symmetry at the Lagrangian level,
either by explicitly picking different gluon states (as in SDYM) or by
picking states that only restore Bose symmetry in the final amplitude
(as in NLSM).  All of these theories can be encapsulated in the
semi-abelian YM theory presented in \cref{eq:semiYM} which obscures
the Bose symmetry of the external states due to the presence of the
two gauge fields $A_\mu$ and $\bar{A}_\mu$.

In addition to the natural tree-level construction, we found that kinetic mixing
between the two gauge fields permits globally color-dual one-loop
integrands, but special care must be taken so that the integrands
produced from semi-abelian YM correspond to SDYM or NLSM.  Due to the
kinetic mixing that stems from the broken Bose symmetry, it is
impossible to construct a two-loop integrand for semi-abelian YM theory,
that is, the theory is one-loop exact.  Obtaining a globally color-dual
integrand beyond one-loop then requires sacrificing more than just Bose
symmetry and in this paper we have argued that off-shell locality must
be abandoned. \footnote{An example of this can already be found in the literature, where Locality is relaxed in \cite{Mogull:2015adi} in
  order to obtain color-dual two-loop integrands with on-shell 4D states.}

We concede building loop-level numerators from rational functions of the kinematics
is rather unnatural from the perspective of point-like quantum field
theories. After all, operators that produce rational functions are typically non-local in their construction. The archetypal example of non-local quantum operators are string vertex operators with $\alpha'$ corrections. These operators promote local tree-level amplitudes to stringy extended objects by integrating over the open string world-sheet:
\be
\mathcal{A}^{\text{YM}} \stackrel{\alpha'}{\longrightarrow} \mathcal{A}^{\text{OS}}t_8 F^4 \int_0^1 dx \,(1-x)^{\alpha' s_{12}-1} x^{\alpha' s_{12}-1}
\ee
The resulting Veneziano factor of Gamma functions produces rational functions of kinematics, while preserving the color-dual structure at in the color-ordering of the Park-Taylor factor in the denominator. While introducing a string world sheet is a jump from what we have found in this work, our funding certainly suggest that we must do \textit{something} to relax the constraint of
\textit{off-shell locality}, and realize locality only in the limit of
\textit{on-shell} kinematics. This could be acheived by either modifying the kinetic term or potentially something more exotic. Below we provide an overview of what one
might expect for such a construction, and provide some simple examples
at tree-level that manifest this structure.

\subsection{Non-local construction of scattering amplitudes}
\label{nonLocalScattering}
As an exemplar of an off-shell non-local construction, consider the simple four-point example of a color-dual
representation of Yang-Mills theory. We can define a functional numerator, $N_{(12|34)}^{\text{YM}}$, as follows, 
\begin{equation}
N_{(12|34)}^{\text{YM}} = \frac{t_8F^4}{3} \frac{s_{13}-s_{12}}{s_{23}{s_{13}}}
\end{equation}
where $N_s = N_{(12|34)}$, $N_t = N_{(14|23)}$ and $N_u = N_{(13|42)}$. By construction, this functionally symmetric numerator is
anti-symmetric and obeys the Jacobi identity:
\begin{equation}\label{nonLocalNum}
N_{(12|34)}^{\text{YM}} + N_{(13|42)}^{\text{YM}} +N_{(14|23)}^{\text{YM}} = 0
\end{equation}
However, it must also factorize to kinematics that are consistent with
the local Feynman rules of \cref{eq:ym-lag}. Evaluating the residue on
the $s=s_{12} \to 0$ pole we find:
\begin{align}
\text{Res}\left[\frac{ N_{s}^{\text{YM}} }{s}\right]^{s=0}&= \frac{2}{3}A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\\
\text{Res}\left[\frac{N_{t}^{\text{YM}} }{t}\right]^{s=0}&= -\frac{1}{3}A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\\
\text{Res}\left[\frac{N_{u}^{\text{YM}} }{u}\right]^{s=0}&= -\frac{1}{3} A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} 
\end{align}
with $t=s_{23}$ and $u=s_{13}$.
Plugging these into a cubic graph representation of Yang-Mills, the
tree level version of \cref{eq:gen-amp}, we find:
\begin{equation}
  \text{Res}\left[\mathcal{A}_4^{\text{YM}}\right]^{s=0}
  = \frac{1}{3} A_{(12|l)}^{\text{YM}}A_{(-l|34)}^{\text{YM}} (2c_s-c_t-c_u)
  =  \mathcal{A}_{(12|l)}^{\text{YM}}\mathcal{A}_{(-l|34)}^{\text{YM}} 
\end{equation}
where we have applied the color structure Jacobi identity,
$c_s+c_t+c_u=0$. In light of our findings in \cref{chefsKiss}, building color-dual numerators in this way where locality is only realized on-shell might be a more natural approach. While we can absolutely apply a generalized gauge
transformation \cite{BCJ} to restore field theoretic locality to the
cubic numerators, this could merely be an aesthetic choice that is only permissible at tree-level.

Moreover, by abandoning traditional notions of off-shell
locality, we incidentally have gained enough functional freedom to massage the
color-dual numerators into a form that is manifestly gauge invariant
for all particles. As an organizational principle, pulling out overall
gauge-invariant factors comes with the added advantage of simplifying
a loop-level construction of color-dual numerators for Yang-Mills.

\subsection{Future Directions}\label{sec:Discussion}
With an eye towards generalizing this construction for future multi-loop studies, we note that the four-point half-ladder secretly makes use of the color-dual structure of NLSM. At four-point, we can construct permutation invariants from the single BCJ basis amplitude of both NLSM and pure YM theory as follows:
\be
s t A^{\text{YM}}_{(s,t)} = t_8F^4 \qquad s t A^{\text{NLSM}}_{(s,t)} = s_{12}s_{23}s_{13}
\ee
Using this, we can redefine our non-local numerator of \cref{nonLocalNum} in a form that recasts color-dual structure to NLSM, and removes all the vector structure of Yang-Mills as an overall $t_8F^4$ factor:
\be
N_{(12|34)}^{\text{YM}} \equiv t_8F^4 \,\frac{N_{(12|34)}^{\text{NLSM}}}{s_{12}s_{23}s_{13}} 
\ee
One advantage of this construction is that reduces some of the $D$-dimensional complexity vector theories that arises due to the mixing between external polarization and
internal loop momenta. But maybe more importantly, it puts all the heavy lifting of functional jacobi relations on the scalar kinematic numerator, $N_{(12|34)}^{\text{NLSM}}$. Thus, rather
than building an ansatz from the irreducible scalar products of
\cref{eq:ym-basis}, one might instead consider the following construction
\begin{equation}
N^{\text{YM}}_{m,L} = \sum_i \mathcal{O}_i \mathcal{R}^{(i)}_{m,L}
\end{equation}
where $\mathcal{O}_i$ are the gauge invariant tensor basis elements of
\cite{Bern:2017tuc,Carrasco:2019yyn}, and $ \mathcal{R}^{(i)}_{m,L}$ are
rational functions of irreducible scalar products of \cref{eq:scalar-basis}. It would be interesting if a similar construction as our toy tree-level example could be uplifted to two-loop using the globally color-dual NLSM integrand that we have computed in this work. We see this as a natural future direction worth investigating that is made possible given the results of this work.
